---
title: "Functional Analysis"
output:
  html_document:
    df_print: paged
    toc: true
    toc_float: true
---

Copyright © 2024 Daniel Zimmermann. <daniel.zimmermann.dz@outlook.com>

You may use, distribute and modify this code under the terms of the MIT license.

You should have received a copy of the MIT license with this file. If not, please visit: https://opensource.org/license/mit

# Setup

```{r setup}

library(tidyverse)
library(clusterProfiler)
library(org.Hs.eg.db)
library(enrichplot)
library(pathview)
library(DOSE)

knitr::opts_chunk$set(fig.width = 10, out.width = "100%", dpi = 300)

dir.create(path = "./results/FA", showWarnings = FALSE)

```

# Results

## MCM1G vs. nHEM

### Gene ontology

#### Biological Process

```{r MCM1GvsHEM_import, warning=FALSE}
out.dir <- "./results/FA/MCM1G_vs_nHEM/"

dir.create(out.dir)

res.1GvsHEM <- read.csv("results/DESeq/MCM1G_vs_nHEM_unfiltered.csv") %>%
  as_tibble() %>%
  arrange(desc(log2FoldChange))

gene_list.1GvsHEM <- res.1GvsHEM$log2FoldChange
names(gene_list.1GvsHEM) <- res.1GvsHEM$ENSEMBL

```


```{r MCM1GvsHEM_go_bp, warning=FALSE}

gseGO.1GvsHEM.BP <- gseGO(gene_list.1GvsHEM,
                          ont = "BP",
                          OrgDb = org.Hs.eg.db,
                          keyType = "ENSEMBL",
                          maxGSSize = 500,
                          minGSSize = 15,
                          eps = 0)

gseGO.1GvsHEM.BP <- simplify(gseGO.1GvsHEM.BP)
gseGO.1GvsHEM.BP <- setReadable(gseGO.1GvsHEM.BP, "org.Hs.eg.db", "ENSEMBL")
gseGO.1GvsHEM.BP

write.csv(data.frame(gseGO.1GvsHEM.BP), file = paste0(out.dir, "gsea_go_bp.csv"), row.names = F)

MCM1GvsHEM.go_bp.plot <- nrow(gseGO.1GvsHEM.BP) > 0

```

```{r MCM1GvsHEM_go_bp_dotplot, eval=MCM1GvsHEM.go_bp.plot, fig.asp=1.75}

dotplot(gseGO.1GvsHEM.BP, showCategory = 15, size = NULL, split = ".sign", font.size = 10) + 
  facet_grid(~.sign) +
  scale_fill_gradient(low = "red", high = "white", limits = c(0, 0.05), name = "adj. p-value")

```

```{r MCM1GvsHEM_go_bp_netplot, eval=MCM1GvsHEM.go_bp.plot, fig.asp=1}

cnetplot(gseGO.1GvsHEM.BP, showCategory = 5, foldChange = gene_list.1GvsHEM, node_label = "category") +
  scale_color_gradient2(low = "blue", mid = "white", high = "red", limits = c(-2, 2), name = "Fold Change", oob = scales::squish)

```

```{r MCM1GvsHEM_go_bp_mapplot, eval=MCM1GvsHEM.go_bp.plot, fig.asp=1}
gseGO.1GvsHEM.BP.pwts <- pairwise_termsim(gseGO.1GvsHEM.BP)

emapplot(gseGO.1GvsHEM.BP.pwts, size_category = 0.7) +
  scale_color_gradient(low = "red", high = "white", limits = c(0, 0.05), name = "adj. p-value")
```

#### Cellular Component

```{r MCM1GvsHEM_go_cc, warning=FALSE}
  

gseGO.1GvsHEM.CC <- gseGO(gene_list.1GvsHEM,
                          ont = "CC",
                          OrgDb = org.Hs.eg.db,
                          keyType = "ENSEMBL",
                          maxGSSize = 500,
                          minGSSize = 15, 
                          eps = 0)

gseGO.1GvsHEM.CC <- simplify(gseGO.1GvsHEM.CC)
gseGO.1GvsHEM.CC <- setReadable(gseGO.1GvsHEM.CC, "org.Hs.eg.db", "ENSEMBL")
gseGO.1GvsHEM.CC

write.csv(data.frame(gseGO.1GvsHEM.CC), file = paste0(out.dir, "gsea_go_cc.csv"), row.names = F)

MCM1GvsHEM.go_cc.plot <- nrow(gseGO.1GvsHEM.CC) > 0

```

```{r MCM1GvsHEM_go_cc_dotplot, eval=MCM1GvsHEM.go_cc.plot, fig.asp=1}

dotplot(gseGO.1GvsHEM.CC, showCategory = 15, size = NULL, split = ".sign", font.size = 10) + 
  facet_grid(~.sign) +
  scale_fill_gradient(low = "red", high = "white", limits = c(0, 0.05), name = "adj. p-value")

```

```{r MCM1GvsHEM_go_cc_netplot, eval=MCM1GvsHEM.go_cc.plot, fig.asp=1}

cnetplot(gseGO.1GvsHEM.CC, showCategory = 5, foldChange = gene_list.1GvsHEM, node_label = "category") +
  scale_color_gradient2(low = "blue", mid = "white", high = "red", limits = c(-2, 2), name = "Fold Change", oob = scales::squish)

```

```{r MCM1GvsHEM_go_cc_mapplot, eval=MCM1GvsHEM.go_cc.plot, fig.asp=1}
gseGO.1GvsHEM.CC.pwts <- pairwise_termsim(gseGO.1GvsHEM.CC)

emapplot(gseGO.1GvsHEM.CC.pwts, size_category = 0.7) +
  scale_color_gradient(low = "red", high = "white", limits = c(0, 0.05), name = "adj. p-value")
```

#### Molecular Function

```{r MCM1GvsHEM_go_mf, warning=FALSE}
  

gseGO.1GvsHEM.MF <- gseGO(gene_list.1GvsHEM,
                          ont = "MF",
                          OrgDb = org.Hs.eg.db,
                          keyType = "ENSEMBL",
                          maxGSSize = 500,
                          minGSSize = 15, 
                          eps = 0)

gseGO.1GvsHEM.MF <- simplify(gseGO.1GvsHEM.MF)
gseGO.1GvsHEM.MF <- setReadable(gseGO.1GvsHEM.MF, "org.Hs.eg.db", "ENSEMBL")
gseGO.1GvsHEM.MF

write.csv(data.frame(gseGO.1GvsHEM.MF), file = paste0(out.dir, "gsea_go_mf.csv"), row.names = F)

MCM1GvsHEM.go_mf.plot <- nrow(gseGO.1GvsHEM.MF) > 0

```

```{r MCM1GvsHEM_go_mf_dotplot, eval=MCM1GvsHEM.go_mf.plot, fig.asp=0.9}

dotplot(gseGO.1GvsHEM.MF, showCategory = 15, size = NULL, split = ".sign", font.size = 10) + 
  facet_grid(~.sign) +
  scale_fill_gradient(low = "red", high = "white", limits = c(0, 0.05), name = "adj. p-value")

```

```{r MCM1GvsHEM_go_mf_netplot, eval=MCM1GvsHEM.go_mf.plot, fig.asp=1}

cnetplot(gseGO.1GvsHEM.MF, showCategory = 5, foldChange = gene_list.1GvsHEM, node_label = "category") +
  scale_color_gradient2(low = "blue", mid = "white", high = "red", limits = c(-2, 2), name = "Fold Change", oob = scales::squish)

```

```{r MCM1GvsHEM_go_mf_mapplot, eval=MCM1GvsHEM.go_mf.plot, fig.asp=1}
gseGO.1GvsHEM.MF.pwts <- pairwise_termsim(gseGO.1GvsHEM.MF)

emapplot(gseGO.1GvsHEM.MF.pwts, size_category = 0.7) +
  scale_color_gradient(low = "red", high = "white", limits = c(0, 0.05), name = "adj. p-value")
```

### KEGG (Kyoto Encyclopedia of Genes and Genomes)

```{r MCM1GvsHEM_translate_ids, warning=FALSE}
ids <- bitr(names(gene_list.1GvsHEM), fromType = "ENSEMBL", toType = "ENTREZID", OrgDb = org.Hs.eg.db)
ids <- ids[!duplicated(ids$ENSEMBL),]

gene_list.1GvsHEM.entrez <- gene_list.1GvsHEM[names(gene_list.1GvsHEM) %in% ids$ENSEMBL]
names(gene_list.1GvsHEM.entrez) <- ids$ENTREZID
gene_list.1GvsHEM.entrez <- gene_list.1GvsHEM.entrez[!duplicated(names(gene_list.1GvsHEM.entrez))]

```


```{r MCM1GvsHEM_kegg, warning=FALSE}

gseKEGG.1GvsHEM <- gseKEGG(gene_list.1GvsHEM.entrez,
                         organism = "hsa",
                         keyType = "ncbi-geneid",
                         maxGSSize = 500,
                         minGSSize = 15, 
                         eps = 0)

gseKEGG.1GvsHEM <- setReadable(gseKEGG.1GvsHEM, "org.Hs.eg.db", "ENTREZID")
gseKEGG.1GvsHEM

write.csv(data.frame(gseKEGG.1GvsHEM), file = paste0(out.dir, "gsea_kegg.csv"), row.names = F)

MCM1GvsHEM.kegg.plot <- nrow(gseKEGG.1GvsHEM) > 0

```


```{r MCM1GvsHEM_kegg_dotplot, eval=MCM1GvsHEM.kegg.plot, fig.asp=0.9}

dotplot(gseKEGG.1GvsHEM, showCategory = 15, size = NULL, split = ".sign", font.size = 9, label_format = 40) + 
  facet_grid(~.sign) +
  scale_fill_gradient(low = "red", high = "white", limits = c(0, 0.05), name = "adj. p-value")

```

```{r MCM1GvsHEM_kegg_netplot, eval=MCM1GvsHEM.kegg.plot, fig.asp=1}

cnetplot(gseKEGG.1GvsHEM, showCategory = 5, foldChange = gene_list.1GvsHEM.entrez, node_label = "category") +
  scale_color_gradient2(low = "blue", mid = "white", high = "red", limits = c(-2, 2), name = "Fold Change", oob = scales::squish)

```

```{r MCM1GvsHEM_kegg_mapplot, eval=MCM1GvsHEM.kegg.plot, fig.asp=1}
gseKEGG.1GvsHEM.pwts <- pairwise_termsim(gseKEGG.1GvsHEM)

emapplot(gseKEGG.1GvsHEM.pwts, size_category = 0.7) +
  scale_color_gradient(low = "red", high = "white", limits = c(0, 0.05), name = "adj. p-value")

```

```{r MCM1GvsHEM_kegg_pathview, eval=MCM1GvsHEM.kegg.plot}

pathway <- "hsa04110"

pathview(gene.data = gene_list.1GvsHEM.entrez, pathway.id = pathway, 
         species = "hsa", low = list(gene = "orangered", cpd = "orangered"),
         mid = list(gene = "white", cpd = "white"), 
         high = list(gene = "deepskyblue", cpd = "deepskyblue"), na.col = "grey")

filename <- paste0(pathway, ".pathview.png")

file.rename(filename, paste0(out.dir, filename))

knitr::include_graphics(paste0(out.dir, filename))

```

### Disease ontology

```{r MCM1GvsHEM_do, warning=FALSE}


gseDO.1GvsHEM <- gseDO(gene_list.1GvsHEM.entrez,
                       maxGSSize = 500,
                       minGSSize = 15, 
                       eps = 0)

gseDO.1GvsHEM <- setReadable(gseDO.1GvsHEM, "org.Hs.eg.db", "ENTREZID")
gseDO.1GvsHEM

write.csv(data.frame(gseDO.1GvsHEM), file = paste0(out.dir, "gsea_do.csv"), row.names = F)

MCM1GvsHEM.DO.plot <- nrow(gseDO.1GvsHEM) > 0

```

```{r MCM1GvsHEM_do_dotplot, eval=MCM1GvsHEM.DO.plot, fig.asp=0.9}

dotplot(gseDO.1GvsHEM, showCategory = 15, size = NULL, split = ".sign", font.size = 9, label_format = 40) + 
  facet_grid(~.sign) +
  scale_fill_gradient(low = "red", high = "white", limits = c(0, 0.05), name = "adj. p-value")

```

```{r MCM1GvsHEM_do_netplot, eval=MCM1GvsHEM.DO.plot, fig.asp=1}

cnetplot(gseDO.1GvsHEM, showCategory = 5, foldChange = gene_list.1GvsHEM.entrez, node_label = "category") +
  scale_color_gradient2(low = "blue", mid = "white", high = "red", limits = c(-2, 2), name = "Fold Change", oob = scales::squish)

```

```{r MCM1GvsHEM_do_mapplot, eval=MCM1GvsHEM.DO.plot, fig.asp=1}

gseDO.1GvsHEM.pwts <- pairwise_termsim(gseDO.1GvsHEM)

emapplot(gseDO.1GvsHEM.pwts, size_category = 0.7) +
  scale_color_gradient(low = "red", high = "white", limits = c(0, 0.05), name = "adj. p-value")

```


## MCMDLN vs. nHEM

### Gene ontology

#### Biological Process

```{r MCMDLNvsHEM_import, warning=FALSE}
out.dir <- "./results/FA/MCMDLN_vs_nHEM/"

dir.create(out.dir)

res.DLNvsHEM <- read.csv("results/DESeq/MCMDLN_vs_nHEM_unfiltered.csv") %>%
  as_tibble() %>%
  arrange(desc(log2FoldChange))

gene_list.DLNvsHEM <- res.DLNvsHEM$log2FoldChange
names(gene_list.DLNvsHEM) <- res.DLNvsHEM$ENSEMBL

```


```{r MCMDLNvsHEM_go_bp, warning=FALSE}


gseGO.DLNvsHEM.BP <- gseGO(gene_list.DLNvsHEM,
                          ont = "BP",
                          OrgDb = org.Hs.eg.db,
                          keyType = "ENSEMBL",
                          maxGSSize = 500,
                          minGSSize = 15, 
                          eps = 0)

gseGO.DLNvsHEM.BP <- simplify(gseGO.DLNvsHEM.BP)
gseGO.DLNvsHEM.BP <- setReadable(gseGO.DLNvsHEM.BP, "org.Hs.eg.db", "ENSEMBL")
gseGO.DLNvsHEM.BP

write.csv(data.frame(gseGO.DLNvsHEM.BP), file = paste0(out.dir, "gsea_go_bp.csv"), row.names = F)

MCMDLNvsHEM.go_bp.plot <- nrow(gseGO.DLNvsHEM.BP) > 0

```

```{r MCMDLNvsHEM_go_bp_dotplot, eval=MCMDLNvsHEM.go_bp.plot, fig.asp=1.25}

dotplot(gseGO.DLNvsHEM.BP, showCategory = 15, size = NULL, split = ".sign", font.size = 10) + 
  facet_grid(~.sign) +
  scale_fill_gradient(low = "red", high = "white", limits = c(0, 0.05), name = "adj. p-value")

```

```{r MCMDLNvsHEM_go_bp_netplot, eval=MCMDLNvsHEM.go_bp.plot, fig.asp=1}

cnetplot(gseGO.DLNvsHEM.BP, showCategory = 5, foldChange = gene_list.DLNvsHEM, node_label = "category") +
  scale_color_gradient2(low = "blue", mid = "white", high = "red", limits = c(-2, 2), name = "Fold Change", oob = scales::squish)

```

```{r MCMDLNvsHEM_go_bp_mapplot, eval=MCMDLNvsHEM.go_bp.plot, fig.asp=1}
gseGO.DLNvsHEM.BP.pwts <- pairwise_termsim(gseGO.DLNvsHEM.BP)

emapplot(gseGO.DLNvsHEM.BP.pwts, size_category = 0.7) +
  scale_color_gradient(low = "red", high = "white", limits = c(0, 0.05), name = "adj. p-value")
```

#### Cellular Component

```{r MCMDLNvsHEM_go_cc, warning=FALSE}
  

gseGO.DLNvsHEM.CC <- gseGO(gene_list.DLNvsHEM,
                          ont = "CC",
                          OrgDb = org.Hs.eg.db,
                          keyType = "ENSEMBL",
                          maxGSSize = 500,
                          minGSSize = 15, 
                          eps = 0)

gseGO.DLNvsHEM.CC <- simplify(gseGO.DLNvsHEM.CC)
gseGO.DLNvsHEM.CC <- setReadable(gseGO.DLNvsHEM.CC, "org.Hs.eg.db", "ENSEMBL")
gseGO.DLNvsHEM.CC

write.csv(data.frame(gseGO.DLNvsHEM.CC), file = paste0(out.dir, "gsea_go_cc.csv"), row.names = F)

MCMDLNvsHEM.go_cc.plot <- nrow(gseGO.DLNvsHEM.CC) > 0

```

```{r MCMDLNvsHEM_go_cc_dotplot, eval=MCMDLNvsHEM.go_cc.plot, fig.asp=0.9}

dotplot(gseGO.DLNvsHEM.CC, showCategory = 15, size = NULL, split = ".sign", font.size = 10) + 
  facet_grid(~.sign) +
  scale_fill_gradient(low = "red", high = "white", limits = c(0, 0.05), name = "adj. p-value")

```

```{r MCMDLNvsHEM_go_cc_netplot, eval=MCMDLNvsHEM.go_cc.plot, fig.asp=1}

cnetplot(gseGO.DLNvsHEM.CC, showCategory = 5, foldChange = gene_list.DLNvsHEM, node_label = "category") +
  scale_color_gradient2(low = "blue", mid = "white", high = "red", limits = c(-2, 2), name = "Fold Change", oob = scales::squish)

```

```{r MCMDLNvsHEM_go_cc_mapplot, eval=MCMDLNvsHEM.go_cc.plot, fig.asp=1}
gseGO.DLNvsHEM.CC.pwts <- pairwise_termsim(gseGO.DLNvsHEM.CC)

emapplot(gseGO.DLNvsHEM.CC.pwts, size_category = 0.7) +
  scale_color_gradient(low = "red", high = "white", limits = c(0, 0.05), name = "adj. p-value")
```

#### Molecular Function

```{r MCMDLNvsHEM_go_mf, warning=FALSE}
  

gseGO.DLNvsHEM.MF <- gseGO(gene_list.DLNvsHEM,
                          ont = "MF",
                          OrgDb = org.Hs.eg.db,
                          keyType = "ENSEMBL",
                          maxGSSize = 500,
                          minGSSize = 15, 
                          eps = 0)

gseGO.DLNvsHEM.MF <- simplify(gseGO.DLNvsHEM.MF)
gseGO.DLNvsHEM.MF <- setReadable(gseGO.DLNvsHEM.MF, "org.Hs.eg.db", "ENSEMBL")
gseGO.DLNvsHEM.MF

write.csv(data.frame(gseGO.DLNvsHEM.MF), file = paste0(out.dir, "gsea_go_mf.csv"), row.names = F)

MCMDLNvsHEM.go_mf.plot <- nrow(gseGO.DLNvsHEM.MF) > 0

```

```{r MCMDLNvsHEM_go_mf_dotplot, eval=MCMDLNvsHEM.go_mf.plot, fig.asp=1.2}

gseGO.DLNvsHEM.MF.plot <- gseGO.DLNvsHEM.MF
gseGO.DLNvsHEM.MF.plot %<>% mutate(Description = ifelse(nchar(Description) <= 75, Description, str_trunc(Description, 75)))

dotplot(gseGO.DLNvsHEM.MF.plot, showCategory = 15, size = NULL, split = ".sign", font.size = 10) + 
  facet_grid(~.sign) +
  scale_fill_gradient(low = "red", high = "white", limits = c(0, 0.05), name = "adj. p-value")

```

```{r MCMDLNvsHEM_go_mf_netplot, eval=MCMDLNvsHEM.go_mf.plot, fig.asp=1}

cnetplot(gseGO.DLNvsHEM.MF, showCategory = 5, foldChange = gene_list.DLNvsHEM, node_label = "category") +
  scale_color_gradient2(low = "blue", mid = "white", high = "red", limits = c(-2, 2), name = "Fold Change", oob = scales::squish)

```

```{r MCMDLNvsHEM_go_mf_mapplot, eval=MCMDLNvsHEM.go_mf.plot, fig.asp=1}
gseGO.DLNvsHEM.MF.pwts <- pairwise_termsim(gseGO.DLNvsHEM.MF)

emapplot(gseGO.DLNvsHEM.MF.pwts, size_category = 0.7) +
  scale_color_gradient(low = "red", high = "white", limits = c(0, 0.05), name = "adj. p-value")
```

### KEGG (Kyoto Encyclopedia of Genes and Genomes)

```{r MCMDLNvsHEM_translate_ids, warning=FALSE}
ids <- bitr(names(gene_list.DLNvsHEM), fromType = "ENSEMBL", toType = "ENTREZID", OrgDb = org.Hs.eg.db)
ids <- ids[!duplicated(ids$ENSEMBL),]

gene_list.DLNvsHEM.entrez <- gene_list.DLNvsHEM[names(gene_list.DLNvsHEM) %in% ids$ENSEMBL]
names(gene_list.DLNvsHEM.entrez) <- ids$ENTREZID
gene_list.DLNvsHEM.entrez <- gene_list.DLNvsHEM.entrez[!duplicated(names(gene_list.DLNvsHEM.entrez))]

```


```{r MCMDLNvsHEM_kegg, warning=FALSE}


gseKEGG.DLNvsHEM <- gseKEGG(gene_list.DLNvsHEM.entrez,
                         organism = "hsa",
                         keyType = "ncbi-geneid",
                         maxGSSize = 500,
                         minGSSize = 15, 
                         eps = 0)

gseKEGG.DLNvsHEM <- setReadable(gseKEGG.DLNvsHEM, "org.Hs.eg.db", "ENTREZID")
gseKEGG.DLNvsHEM

write.csv(data.frame(gseKEGG.DLNvsHEM), file = paste0(out.dir, "gsea_kegg.csv"), row.names = F)

MCMDLNvsHEM.kegg.plot <- nrow(gseKEGG.DLNvsHEM) > 0

```


```{r MCMDLNvsHEM_kegg_dotplot, eval=MCMDLNvsHEM.kegg.plot, fig.asp=0.9}

dotplot(gseKEGG.DLNvsHEM, showCategory = 15, size = NULL, split = ".sign", font.size = 9, label_format = 40) + 
  facet_grid(~.sign) +
  scale_fill_gradient(low = "red", high = "white", limits = c(0, 0.05), name = "adj. p-value")

```

```{r MCMDLNvsHEM_kegg_netplot, eval=MCMDLNvsHEM.kegg.plot, fig.asp=1}

cnetplot(gseKEGG.DLNvsHEM, showCategory = 5, foldChange = gene_list.DLNvsHEM.entrez, node_label = "category") +
  scale_color_gradient2(low = "blue", mid = "white", high = "red", limits = c(-2, 2), name = "Fold Change", oob = scales::squish)

```

```{r MCMDLNvsHEM_kegg_mapplot, eval=MCMDLNvsHEM.kegg.plot, fig.asp=1}
gseKEGG.DLNvsHEM.pwts <- pairwise_termsim(gseKEGG.DLNvsHEM)

emapplot(gseKEGG.DLNvsHEM.pwts, size_category = 0.7) +
  scale_color_gradient(low = "red", high = "white", limits = c(0, 0.05), name = "adj. p-value")

```

```{r MCMDLNvsHEM_kegg_pathview, eval=MCMDLNvsHEM.kegg.plot}

pathway <- "hsa04110"

pathview(gene.data = gene_list.DLNvsHEM.entrez, pathway.id = pathway, 
         species = "hsa", low = list(gene = "orangered", cpd = "orangered"),
         mid = list(gene = "white", cpd = "white"), 
         high = list(gene = "deepskyblue", cpd = "deepskyblue"), na.col = "grey")

filename <- paste0(pathway, ".pathview.png")

file.rename(filename, paste0(out.dir, filename))

knitr::include_graphics(paste0(out.dir, filename))

```

### Disease ontology

```{r MCMDLNvsHEM_do, warning=FALSE}


gseDO.DLNvsHEM <- gseDO(gene_list.DLNvsHEM.entrez,
                     maxGSSize = 500,
                     minGSSize = 15, 
                     eps = 0)

gseDO.DLNvsHEM <- setReadable(gseDO.DLNvsHEM, "org.Hs.eg.db", "ENTREZID")
gseDO.DLNvsHEM

write.csv(data.frame(gseDO.DLNvsHEM), file = paste0(out.dir, "gsea_do.csv"), row.names = F)

MCMDLNvsHEM.DO.plot <- nrow(gseDO.DLNvsHEM) > 0

```

```{r MCMDLNvsHEM_do_dotplot, eval=MCMDLNvsHEM.DO.plot, fig.asp=0.9}

dotplot(gseDO.DLNvsHEM, showCategory = 15, size = NULL, split = ".sign", font.size = 9, label_format = 40) + 
  facet_grid(~.sign) +
  scale_fill_gradient(low = "red", high = "white", limits = c(0, 0.05), name = "adj. p-value")

```

```{r MCMDLNvsHEM_do_netplot, eval=MCMDLNvsHEM.DO.plot, fig.asp=1}

cnetplot(gseDO.DLNvsHEM, showCategory = 5, foldChange = gene_list.DLNvsHEM.entrez, node_label = "category") +
  scale_color_gradient2(low = "blue", mid = "white", high = "red", limits = c(-2, 2), name = "Fold Change", oob = scales::squish)

```

```{r MCMDLNvsHEM_do_mapplot, eval=MCMDLNvsHEM.DO.plot, fig.asp=1}

gseDO.DLNvsHEM.pwts <- pairwise_termsim(gseDO.DLNvsHEM)

emapplot(gseDO.DLNvsHEM.pwts, size_category = 0.7) +
  scale_color_gradient(low = "red", high = "white", limits = c(0, 0.05), name = "adj. p-value")

```


## MCMDLN vs. MCM1G

### Gene ontology

#### Biological Process

```{r MCMDLNvsMCM1G_import, warning=FALSE}
out.dir <- "./results/FA/MCMDLN_vs_MCM1G/"

dir.create(out.dir)

res.DLNvs1G <- read.csv("results/DESeq/MCMDLN_vs_MCM1G_unfiltered.csv") %>%
  as_tibble() %>%
  arrange(desc(log2FoldChange))

gene_list.DLNvs1G <- res.DLNvs1G$log2FoldChange
names(gene_list.DLNvs1G) <- res.DLNvs1G$ENSEMBL

```


```{r MCMDLNvsMCM1G_go_bp, warning=FALSE}


gseGO.DLNvs1G.BP <- gseGO(gene_list.DLNvs1G,
                          ont = "BP",
                          OrgDb = org.Hs.eg.db,
                          keyType = "ENSEMBL",
                          maxGSSize = 500,
                          minGSSize = 15, 
                          eps = 0)

gseGO.DLNvs1G.BP <- simplify(gseGO.DLNvs1G.BP)
gseGO.DLNvs1G.BP <- setReadable(gseGO.DLNvs1G.BP, "org.Hs.eg.db", "ENSEMBL")
gseGO.DLNvs1G.BP

write.csv(data.frame(gseGO.DLNvs1G.BP), file = paste0(out.dir, "gsea_go_bp.csv"), row.names = F)

MCMDLNvsMCM1G.go_bp.plot <- nrow(gseGO.DLNvs1G.BP) > 0

```

```{r MCMDLNvsMCM1G_go_bp_dotplot, eval=MCMDLNvsMCM1G.go_bp.plot, fig.asp=1.25}

dotplot(gseGO.DLNvs1G.BP, showCategory = 15, size = NULL, split = ".sign", font.size = 10) + 
  facet_grid(~.sign) +
  scale_fill_gradient(low = "red", high = "white", limits = c(0, 0.05), name = "adj. p-value")

```

```{r MCMDLNvsMCM1G_go_bp_netplot, eval=MCMDLNvsMCM1G.go_bp.plot, fig.asp=1.2}

cnetplot(gseGO.DLNvs1G.BP, showCategory = 5, foldChange = gene_list.DLNvs1G, node_label = "category") +
  scale_color_gradient2(low = "blue", mid = "white", high = "red", limits = c(-2, 2), name = "Fold Change", oob = scales::squish)

```

```{r MCMDLNvsMCM1G_go_bp_mapplot, eval=MCMDLNvsMCM1G.go_bp.plot, fig.asp=1}
gseGO.DLNvs1G.BP.pwts <- pairwise_termsim(gseGO.DLNvs1G.BP)

emapplot(gseGO.DLNvs1G.BP.pwts, size_category = 0.7) +
  scale_color_gradient(low = "red", high = "white", limits = c(0, 0.05), name = "adj. p-value")
```

#### Cellular Component

```{r MCMDLNvsMCM1G_go_cc, warning=FALSE}
  

gseGO.DLNvs1G.CC <- gseGO(gene_list.DLNvs1G,
                          ont = "CC",
                          OrgDb = org.Hs.eg.db,
                          keyType = "ENSEMBL",
                          maxGSSize = 500,
                          minGSSize = 15, 
                          eps = 0)

gseGO.DLNvs1G.CC <- simplify(gseGO.DLNvs1G.CC)
gseGO.DLNvs1G.CC <- setReadable(gseGO.DLNvs1G.CC, "org.Hs.eg.db", "ENSEMBL")
gseGO.DLNvs1G.CC

write.csv(data.frame(gseGO.DLNvs1G.CC), file = paste0(out.dir, "gsea_go_cc.csv"), row.names = F)

MCMDLNvsMCM1G.go_cc.plot <- nrow(gseGO.DLNvs1G.CC) > 0

```

```{r MCMDLNvsMCM1G_go_cc_dotplot, eval=MCMDLNvsMCM1G.go_cc.plot, fig.asp=0.9}

dotplot(gseGO.DLNvs1G.CC, showCategory = 15, size = NULL, split = ".sign", font.size = 10) + 
  facet_grid(~.sign) +
  scale_fill_gradient(low = "red", high = "white", limits = c(0, 0.05), name = "adj. p-value")

```

```{r MCMDLNvsMCM1G_go_cc_netplot, eval=MCMDLNvsMCM1G.go_cc.plot, fig.asp=1.2}

cnetplot(gseGO.DLNvs1G.CC, showCategory = 5, foldChange = gene_list.DLNvs1G, node_label = "category") +
  scale_color_gradient2(low = "blue", mid = "white", high = "red", limits = c(-2, 2), name = "Fold Change", oob = scales::squish)

```

```{r MCMDLNvsMCM1G_go_cc_mapplot, eval=MCMDLNvsMCM1G.go_cc.plot, fig.asp=1}
gseGO.DLNvs1G.CC.pwts <- pairwise_termsim(gseGO.DLNvs1G.CC)

emapplot(gseGO.DLNvs1G.CC.pwts, size_category = 0.7) +
  scale_color_gradient(low = "red", high = "white", limits = c(0, 0.05), name = "adj. p-value")
```

#### Molecular Function

```{r MCMDLNvsMCM1G_go_mf, warning=FALSE}
  

gseGO.DLNvs1G.MF <- gseGO(gene_list.DLNvs1G,
                          ont = "MF",
                          OrgDb = org.Hs.eg.db,
                          keyType = "ENSEMBL",
                          maxGSSize = 500,
                          minGSSize = 15, 
                          eps = 0)

gseGO.DLNvs1G.MF <- simplify(gseGO.DLNvs1G.MF)
gseGO.DLNvs1G.MF <- setReadable(gseGO.DLNvs1G.MF, "org.Hs.eg.db", "ENSEMBL")
gseGO.DLNvs1G.MF

write.csv(data.frame(gseGO.DLNvs1G.MF), file = paste0(out.dir, "gsea_go_mf.csv"), row.names = F)

MCMDLNvsMCM1G.go_mf.plot <- nrow(gseGO.DLNvs1G.MF) > 0

```

```{r MCMDLNvsMCM1G_go_mf_dotplot, eval=MCMDLNvsMCM1G.go_mf.plot, fig.asp=0.5}

dotplot(gseGO.DLNvs1G.MF, showCategory = 15, size = NULL, split = ".sign", font.size = 10) + 
  facet_grid(~.sign) +
  scale_fill_gradient(low = "red", high = "white", limits = c(0, 0.05), name = "adj. p-value")

```

```{r MCMDLNvsMCM1G_go_mf_netplot, eval=MCMDLNvsMCM1G.go_mf.plot, fig.asp=1}

cnetplot(gseGO.DLNvs1G.MF, showCategory = 5, foldChange = gene_list.DLNvs1G, node_label = "category") +
  scale_color_gradient2(low = "blue", mid = "white", high = "red", limits = c(-2, 2), name = "Fold Change", oob = scales::squish)

```

```{r MCMDLNvsMCM1G_go_mf_mapplot, eval=MCMDLNvsMCM1G.go_mf.plot, fig.asp=1}
gseGO.DLNvs1G.MF.pwts <- pairwise_termsim(gseGO.DLNvs1G.MF)

emapplot(gseGO.DLNvs1G.MF.pwts, size_category = 0.7) +
  scale_color_gradient(low = "red", high = "white", limits = c(0, 0.05), name = "adj. p-value")
```

### KEGG (Kyoto Encyclopedia of Genes and Genomes)

```{r MCMDLNvsMCM1G_translate_ids, warning=FALSE}
ids <- bitr(names(gene_list.DLNvs1G), fromType = "ENSEMBL", toType = "ENTREZID", OrgDb = org.Hs.eg.db)
ids <- ids[!duplicated(ids$ENSEMBL),]

gene_list.DLNvs1G.entrez <- gene_list.DLNvs1G[names(gene_list.DLNvs1G) %in% ids$ENSEMBL]
names(gene_list.DLNvs1G.entrez) <- ids$ENTREZID
gene_list.DLNvs1G.entrez <- gene_list.DLNvs1G.entrez[!duplicated(names(gene_list.DLNvs1G.entrez))]

```


```{r MCMDLNvsMCM1G_kegg, warning=FALSE}


gseKEGG.DLNvs1G <- gseKEGG(gene_list.DLNvs1G.entrez,
                         organism = "hsa",
                         keyType = "ncbi-geneid",
                         maxGSSize = 500,
                         minGSSize = 15,
                         eps = 0)

gseKEGG.DLNvs1G <- setReadable(gseKEGG.DLNvs1G, "org.Hs.eg.db", "ENTREZID")
gseKEGG.DLNvs1G

write.csv(data.frame(gseKEGG.DLNvs1G), file = paste0(out.dir, "gsea_kegg.csv"), row.names = F)

MCMDLNvsMCM1G.kegg.plot <- nrow(gseKEGG.DLNvs1G) > 0

```


```{r MCMDLNvsMCM1G_kegg_dotplot, eval=MCMDLNvsMCM1G.kegg.plot, fig.asp=0.6}

dotplot(gseKEGG.DLNvs1G, showCategory = 15, size = NULL, split = ".sign", font.size = 9, label_format = 40) + 
  facet_grid(~.sign) +
  scale_fill_gradient(low = "red", high = "white", limits = c(0, 0.05), name = "adj. p-value")

```

```{r MCMDLNvsMCM1G_kegg_netplot, eval=MCMDLNvsMCM1G.kegg.plot, fig.asp=1}

cnetplot(gseKEGG.DLNvs1G, showCategory = 5, foldChange = gene_list.DLNvs1G.entrez, node_label = "category") +
  scale_color_gradient2(low = "blue", mid = "white", high = "red", limits = c(-2, 2), name = "Fold Change", oob = scales::squish)

```

```{r MCMDLNvsMCM1G_kegg_mapplot, eval=MCMDLNvsMCM1G.kegg.plot, fig.asp=1}
gseKEGG.DLNvs1G.pwts <- pairwise_termsim(gseKEGG.DLNvs1G)

emapplot(gseKEGG.DLNvs1G.pwts, size_category = 0.7) +
  scale_color_gradient(low = "red", high = "white", limits = c(0, 0.05), name = "adj. p-value")

```

```{r MCMDLNvsMCM1G_kegg_pathview, eval=MCMDLNvsMCM1G.kegg.plot}

pathway <- "hsa04110"

pathview(gene.data = gene_list.DLNvs1G.entrez, pathway.id = pathway, 
         species = "hsa", low = list(gene = "orangered", cpd = "orangered"),
         mid = list(gene = "white", cpd = "white"), 
         high = list(gene = "deepskyblue", cpd = "deepskyblue"), na.col = "grey")

filename <- paste0(pathway, ".pathview.png")

file.rename(filename, paste0(out.dir, filename))

knitr::include_graphics(paste0(out.dir, filename))

```

### Disease ontology

```{r MCMDLNvsMCM1G_do, warning=FALSE}


gseDO.DLNvs1G <- gseDO(gene_list.DLNvs1G.entrez,
                     maxGSSize = 500,
                     minGSSize = 15, 
                     eps = 0)

gseDO.DLNvs1G <- setReadable(gseDO.DLNvs1G, "org.Hs.eg.db", "ENTREZID")
gseDO.DLNvs1G

write.csv(data.frame(gseDO.DLNvs1G), file = paste0(out.dir, "gsea_do.csv"), row.names = F)

MCMDLNvsMCM1G.DO.plot <- nrow(gseDO.DLNvs1G) > 0

```

```{r MCMDLNvsMCM1G_do_dotplot, eval=MCMDLNvsMCM1G.DO.plot, fig.asp=0.6}

dotplot(gseDO.DLNvs1G, showCategory = 15, size = NULL, split = ".sign", font.size = 9, label_format = 40) + 
  facet_grid(~.sign) +
  scale_fill_gradient(low = "red", high = "white", limits = c(0, 0.05), name = "adj. p-value")

```

```{r MCMDLNvsMCM1G_do_netplot, eval=MCMDLNvsMCM1G.DO.plot, fig.asp=1}

cnetplot(gseDO.DLNvs1G, showCategory = 5, foldChange = gene_list.DLNvs1G.entrez, node_label = "category") +
  scale_color_gradient2(low = "blue", mid = "white", high = "red", limits = c(-2, 2), name = "Fold Change", oob = scales::squish)

```

```{r MCMDLNvsMCM1G_do_mapplot, eval=MCMDLNvsMCM1G.DO.plot, fig.asp=1}

gseDO.DLNvs1G.pwts <- pairwise_termsim(gseDO.DLNvs1G)

emapplot(gseDO.DLNvs1G.pwts, size_category = 0.7) +
  scale_color_gradient(low = "red", high = "white", limits = c(0, 0.05), name = "adj. p-value")

```


## MCMDLN3D vs. MCMDLN

### Gene ontology

#### Biological Process

```{r MCMDLN3DvsMCMDLN_import, warning=FALSE}
out.dir <- "./results/FA/MCMDLN3D_vs_MCMDLN/"

dir.create(out.dir)

res.DLN3DvsDLN <- read.csv("results/DESeq/MCMDLN3D_vs_MCMDLN_unfiltered.csv") %>%
  as_tibble() %>%
  arrange(desc(log2FoldChange))

gene_list.DLN3DvsDLN <- res.DLN3DvsDLN$log2FoldChange
names(gene_list.DLN3DvsDLN) <- res.DLN3DvsDLN$ENSEMBL

```


```{r MCMDLN3DvsMCMDLN_go_bp, warning=FALSE}


gseGO.DLN3DvsDLN.BP <- gseGO(gene_list.DLN3DvsDLN,
                          ont = "BP",
                          OrgDb = org.Hs.eg.db,
                          keyType = "ENSEMBL",
                          maxGSSize = 500,
                          minGSSize = 15, 
                          eps = 0)

gseGO.DLN3DvsDLN.BP <- simplify(gseGO.DLN3DvsDLN.BP)
gseGO.DLN3DvsDLN.BP <- setReadable(gseGO.DLN3DvsDLN.BP, "org.Hs.eg.db", "ENSEMBL")
gseGO.DLN3DvsDLN.BP

write.csv(data.frame(gseGO.DLN3DvsDLN.BP), file = paste0(out.dir, "gsea_go_bp.csv"), row.names = F)

MCMDLN3DvsMCMDLN.go_bp.plot <- nrow(gseGO.DLN3DvsDLN.BP) > 0

```

```{r MCMDLN3DvsMCMDLN_go_bp_dotplot, eval=MCMDLN3DvsMCMDLN.go_bp.plot, fig.asp=1.25}

dotplot(gseGO.DLN3DvsDLN.BP, showCategory = 15, size = NULL, split = ".sign", font.size = 10) + 
  facet_grid(~.sign) +
  scale_fill_gradient(low = "red", high = "white", limits = c(0, 0.05), name = "adj. p-value")

```

```{r MCMDLN3DvsMCMDLN_go_bp_netplot, eval=MCMDLN3DvsMCMDLN.go_bp.plot, fig.asp=1.2}

cnetplot(gseGO.DLN3DvsDLN.BP, showCategory = 5, foldChange = gene_list.DLN3DvsDLN, node_label = "category") +
  scale_color_gradient2(low = "blue", mid = "white", high = "red", limits = c(-2, 2), name = "Fold Change", oob = scales::squish)

```

```{r MCMDLN3DvsMCMDLN_go_bp_mapplot, eval=MCMDLN3DvsMCMDLN.go_bp.plot, fig.asp=1}
gseGO.DLN3DvsDLN.BP.pwts <- pairwise_termsim(gseGO.DLN3DvsDLN.BP)

emapplot(gseGO.DLN3DvsDLN.BP.pwts, size_category = 0.7) +
  scale_color_gradient(low = "red", high = "white", limits = c(0, 0.05), name = "adj. p-value")
```

#### Cellular Component

```{r MCMDLN3DvsMCMDLN_go_cc, warning=FALSE}
  

gseGO.DLN3DvsDLN.CC <- gseGO(gene_list.DLN3DvsDLN,
                          ont = "CC",
                          OrgDb = org.Hs.eg.db,
                          keyType = "ENSEMBL",
                          maxGSSize = 500,
                          minGSSize = 15, 
                          eps = 0)

gseGO.DLN3DvsDLN.CC <- simplify(gseGO.DLN3DvsDLN.CC)
gseGO.DLN3DvsDLN.CC <- setReadable(gseGO.DLN3DvsDLN.CC, "org.Hs.eg.db", "ENSEMBL")
gseGO.DLN3DvsDLN.CC

write.csv(data.frame(gseGO.DLN3DvsDLN.CC), file = paste0(out.dir, "gsea_go_cc.csv"), row.names = F)

MCMDLN3DvsMCMDLN.go_cc.plot <- nrow(gseGO.DLN3DvsDLN.CC) > 0

```

```{r MCMDLN3DvsMCMDLN_go_cc_dotplot, eval=MCMDLN3DvsMCMDLN.go_cc.plot, fig.asp=1}

dotplot(gseGO.DLN3DvsDLN.CC, showCategory = 15, size = NULL, split = ".sign", font.size = 10) + 
  facet_grid(~.sign) +
  scale_fill_gradient(low = "red", high = "white", limits = c(0, 0.05), name = "adj. p-value")

```

```{r MCMDLN3DvsMCMDLN_go_cc_netplot, eval=MCMDLN3DvsMCMDLN.go_cc.plot, fig.asp=1.2}

cnetplot(gseGO.DLN3DvsDLN.CC, showCategory = 5, foldChange = gene_list.DLN3DvsDLN, node_label = "category") +
  scale_color_gradient2(low = "blue", mid = "white", high = "red", limits = c(-2, 2), name = "Fold Change", oob = scales::squish)

```

```{r MCMDLN3DvsMCMDLN_go_cc_mapplot, eval=MCMDLN3DvsMCMDLN.go_cc.plot, fig.asp=1}
gseGO.DLN3DvsDLN.CC.pwts <- pairwise_termsim(gseGO.DLN3DvsDLN.CC)

emapplot(gseGO.DLN3DvsDLN.CC.pwts, size_category = 0.7) +
  scale_color_gradient(low = "red", high = "white", limits = c(0, 0.05), name = "adj. p-value")
```

#### Molecular Function

```{r MCMDLN3DvsMCMDLN_go_mf, warning=FALSE}
  

gseGO.DLN3DvsDLN.MF <- gseGO(gene_list.DLN3DvsDLN,
                          ont = "MF",
                          OrgDb = org.Hs.eg.db,
                          keyType = "ENSEMBL",
                          maxGSSize = 500,
                          minGSSize = 15, 
                          eps = 0)

gseGO.DLN3DvsDLN.MF <- simplify(gseGO.DLN3DvsDLN.MF)
gseGO.DLN3DvsDLN.MF <- setReadable(gseGO.DLN3DvsDLN.MF, "org.Hs.eg.db", "ENSEMBL")
gseGO.DLN3DvsDLN.MF

write.csv(data.frame(gseGO.DLN3DvsDLN.MF), file = paste0(out.dir, "gsea_go_mf.csv"), row.names = F)

MCMDLN3DvsMCMDLN.go_mf.plot <- nrow(gseGO.DLN3DvsDLN.MF) > 0

```

```{r MCMDLN3DvsMCMDLN_go_mf_dotplot, eval=MCMDLN3DvsMCMDLN.go_mf.plot, fig.asp=1.2}

dotplot(gseGO.DLN3DvsDLN.MF, showCategory = 15, size = NULL, split = ".sign", font.size = 10) + 
  facet_grid(~.sign) +
  scale_fill_gradient(low = "red", high = "white", limits = c(0, 0.05), name = "adj. p-value")

```

```{r MCMDLN3DvsMCMDLN_go_mf_netplot, eval=MCMDLN3DvsMCMDLN.go_mf.plot, fig.asp=1}

cnetplot(gseGO.DLN3DvsDLN.MF, showCategory = 5, foldChange = gene_list.DLN3DvsDLN, node_label = "category") +
  scale_color_gradient2(low = "blue", mid = "white", high = "red", limits = c(-2, 2), name = "Fold Change", oob = scales::squish)

```

```{r MCMDLN3DvsMCMDLN_go_mf_mapplot, eval=MCMDLN3DvsMCMDLN.go_mf.plot, fig.asp=1}
gseGO.DLN3DvsDLN.MF.pwts <- pairwise_termsim(gseGO.DLN3DvsDLN.MF)

emapplot(gseGO.DLN3DvsDLN.MF.pwts, size_category = 0.7) +
  scale_color_gradient(low = "red", high = "white", limits = c(0, 0.05), name = "adj. p-value")
```

### KEGG (Kyoto Encyclopedia of Genes and Genomes)

```{r MCMDLN3DvsMCMDLN_translate_ids, warning=FALSE}
ids <- bitr(names(gene_list.DLN3DvsDLN), fromType = "ENSEMBL", toType = "ENTREZID", OrgDb = org.Hs.eg.db)
ids <- ids[!duplicated(ids$ENSEMBL),]

gene_list.DLN3DvsDLN.entrez <- gene_list.DLN3DvsDLN[names(gene_list.DLN3DvsDLN) %in% ids$ENSEMBL]
names(gene_list.DLN3DvsDLN.entrez) <- ids$ENTREZID
gene_list.DLN3DvsDLN.entrez <- gene_list.DLN3DvsDLN.entrez[!duplicated(names(gene_list.DLN3DvsDLN.entrez))]

```


```{r MCMDLN3DvsMCMDLN_kegg, warning=FALSE}


gseKEGG.DLN3DvsDLN <- gseKEGG(gene_list.DLN3DvsDLN.entrez,
                         organism = "hsa",
                         keyType = "ncbi-geneid",
                         maxGSSize = 500,
                         minGSSize = 15, 
                         eps = 0)

gseKEGG.DLN3DvsDLN <- setReadable(gseKEGG.DLN3DvsDLN, "org.Hs.eg.db", "ENTREZID")
gseKEGG.DLN3DvsDLN

write.csv(data.frame(gseKEGG.DLN3DvsDLN), file = paste0(out.dir, "gsea_kegg.csv"), row.names = F)

MCMDLN3DvsMCMDLN.kegg.plot <- nrow(gseKEGG.DLN3DvsDLN) > 0

```


```{r MCMDLN3DvsMCMDLN_kegg_dotplot, eval=MCMDLN3DvsMCMDLN.kegg.plot, fig.asp=0.9}

dotplot(gseKEGG.DLN3DvsDLN, showCategory = 15, size = NULL, split = ".sign", font.size = 9, label_format = 40) + 
  facet_grid(~.sign) +
  scale_fill_gradient(low = "red", high = "white", limits = c(0, 0.05), name = "adj. p-value")

```

```{r MCMDLN3DvsMCMDLN_kegg_netplot, eval=MCMDLN3DvsMCMDLN.kegg.plot, fig.asp=1}

cnetplot(gseKEGG.DLN3DvsDLN, showCategory = 5, foldChange = gene_list.DLN3DvsDLN.entrez, node_label = "category") +
  scale_color_gradient2(low = "blue", mid = "white", high = "red", limits = c(-2, 2), name = "Fold Change", oob = scales::squish)

```

```{r MCMDLN3DvsMCMDLN_kegg_mapplot, eval=MCMDLN3DvsMCMDLN.kegg.plot, fig.asp=1}
gseKEGG.DLN3DvsDLN.pwts <- pairwise_termsim(gseKEGG.DLN3DvsDLN)

emapplot(gseKEGG.DLN3DvsDLN.pwts, size_category = 0.7) +
  scale_color_gradient(low = "red", high = "white", limits = c(0, 0.05), name = "adj. p-value")

```

```{r MCMDLN3DvsMCMDLN_kegg_pathview, eval=MCMDLN3DvsMCMDLN.kegg.plot}

pathway <- "hsa04110"

pathview(gene.data = gene_list.DLN3DvsDLN.entrez, pathway.id = pathway, 
         species = "hsa", low = list(gene = "orangered", cpd = "orangered"),
         mid = list(gene = "white", cpd = "white"), 
         high = list(gene = "deepskyblue", cpd = "deepskyblue"), na.col = "grey")

filename <- paste0(pathway, ".pathview.png")

file.rename(filename, paste0(out.dir, filename))

knitr::include_graphics(paste0(out.dir, filename))

```

### Disease ontology

```{r MCMDLN3DvsMCMDLN_do, warning=FALSE}


gseDO.DLN3DvsDLN <- gseDO(gene_list.DLN3DvsDLN.entrez,
                     maxGSSize = 500,
                     minGSSize = 15, 
                     eps = 0)

gseDO.DLN3DvsDLN <- setReadable(gseDO.DLN3DvsDLN, "org.Hs.eg.db", "ENTREZID")
gseDO.DLN3DvsDLN

write.csv(data.frame(gseDO.DLN3DvsDLN), file = paste0(out.dir, "gsea_do.csv"), row.names = F)

MCMDLN3DvsMCMDLN.DO.plot <- nrow(gseDO.DLN3DvsDLN) > 0

```

```{r MCMDLN3DvsMCMDLN_do_dotplot, eval=MCMDLN3DvsMCMDLN.DO.plot, fig.asp=0.7}

dotplot(gseDO.DLN3DvsDLN, showCategory = 15, size = NULL, split = ".sign", font.size = 9, label_format = 40) + 
  facet_grid(~.sign) +
  scale_fill_gradient(low = "red", high = "white", limits = c(0, 0.05), name = "adj. p-value")

```

```{r MCMDLN3DvsMCMDLN_do_netplot, eval=MCMDLN3DvsMCMDLN.DO.plot, fig.asp=1}

cnetplot(gseDO.DLN3DvsDLN, showCategory = 5, foldChange = gene_list.DLN3DvsDLN.entrez, node_label = "category") +
  scale_color_gradient2(low = "blue", mid = "white", high = "red", limits = c(-2, 2), name = "Fold Change", oob = scales::squish)

```

```{r MCMDLN3DvsMCMDLN_do_mapplot, eval=MCMDLN3DvsMCMDLN.DO.plot, fig.asp=1}

gseDO.DLN3DvsDLN.pwts <- pairwise_termsim(gseDO.DLN3DvsDLN)

emapplot(gseDO.DLN3DvsDLN.pwts, size_category = 0.7) +
  scale_color_gradient(low = "red", high = "white", limits = c(0, 0.05), name = "adj. p-value")

```